The watershed transform in ITK - discussion and new developments

### 2006-04-23 article/Article.tex
Fixed conflicts in article
darcs-hash:20060423232523-fafb9-9348987c8c297faf893ae4c66478b9ffc4695056.gz

\subsubsection{Algorithm 2}
\label{sect:BeucherAlgorithm}

@@

\subsubsection{Algorithm 3}
\label{sect:MeyerAlgorithm}

@@

\section{Timing and performance}

streaming and threading.

Preventing boundary checks - not really the itk way, but if a user is
willing to throw away the image border by treating it as a watershed
line then there is no need to test for boundary conditions. There
isn't even any need to use the face calculator approach because border
pixels will be initially tagged as processed, and therefore we can be
certain their neighbors are never checked.

>>>>

\subsubsection{Algorithm 2 - commonly attributed to Beucher.}
\label{sect:BeucherAlgorithm}

@@

\subsubsection{Algorithm 3 - commonly attributed to Meyer.}
\label{sect:MeyerAlgorithm}

@@

\section{Timing and performance}
The relative performance of the new filters are shown in
Table~\ref{tbl:perf}. The timings were obtained using the {\tt
./perf3D images/ESCells.img}. The times for steps involved in a
morphological watershed are shown in the first few columns, with the
time from a combined filter and the itk watershed shown last. It
should be noted that the itk watershed does not allow connectivity to
be modified by the user - it always uses face connected mode and
doesn't have an inbuilt mechanism for boundary selection. The existing
ITK watershed is therefore somewhat slower than the new implementation
when used to compute a traditional watershed. The new implementations
are not easily threadable and therefore may prove to be slower on some
systems. However, as discussed in earlier sections, the new
implementation produces more stable results than the rainfall model
and can be used in a marker based approach.

\begin{table}[htbp]
\centering
\begin{tabular}{cccccccc}
\hline
FC & MB & Find Min &  Label mins & Watershed from markers & Total & Morph watershed &  itk watershed \\
\hline
0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
\hline
\end{tabular}
\caption{Execution times. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf}}
\end{table}

\subsection{Other performance issues/comments}
\begin{itemize}
\item Preventing boundary checks - not really the itk way, but if a user is
willing to throw away the image border by treating it as a watershed
line then there is no need to test for boundary conditions. There
isn't even any need to use the face calculator approach because border
pixels will be initially tagged as processed, and therefore we can be
certain their neighbors are never checked.
\item Streaming and threading  - queue based algorithms are generally difficult 
to thread efficiently and generally need to have all of the input
available so prevent streaming from working well. The existing ITK
watershed does provide a multithreaded implementation and similar
techniques may be possible with the new implementation. Other
possibilities include using threads to improve performance of queues.
\end{itemize}
